---
title: "Tracking the Stealthy Killer"
date: "06/03/2020"
output: md_document
---

```{r setup, include=FALSE}
# This chunk enables us to avoid a few warnings and output in the document body.
knitr::opts_chunk$set(echo = TRUE) 
library(ggplot2)
library(data.table)

```

This document details the analysis for “Tracking the stealthy killer”.

### Packages and data:
```{r packages, eval = T}

# Load country-china tour group data
pdat <- read.csv("chinese_tourism.csv")

# Load updated covid data
JH <- read.csv("03-04-2020_JH_cases.csv" )

# Merge the two
JH$cases <- ave(JH$Confirmed, JH$Country.Region, FUN = function(x) sum(x))
JH$country <- as.character(JH$Country.Region)
JH$country[JH$country == "US"] <- "United States"
JH$country[JH$country == "UK"] <- "United Kingdom"
pdat$cases <- NULL
pdat <- merge(pdat, unique(JH[, c("country", "cases")]), by = "country", all.x = T)

```

### Source for Chinese tourism data:
* Ministry of Culture and Tourism of the People’s Republic of China (2020). See: http://zwgk.mct.gov.cn/auto255/202001/t20200102_850061.html?keywords=%E5%87%BA%E5%A2%83%E6%97%85%E6%B8%B8

The Ministry of Culture and Tourism track tour group visitors to and from China on a consistent basis across the top 30 recipient/sender countries and all other tour group travels by continent. For countries which were not in the top 30, we distributed the remaining tourists to that continent ('Other Europe' in the data) according to countries' logged population size. 

While not all tourists travel in tour groups, this data was the best recent proxy for total tourism flows that we could identify. In particular, it tracked both incoming and outcoming tourism, was available for a broad set of countries, tracked tourists final destination/origin, and was more recent than other sources (quarter 3 of 2019).

### Source for case data:
* Dong E, Du H, Gardner L. An interactive web-based dashboard to track COVID-19 in real time. Lancet Infect Dis; published online Feb 19. https://doi.org/10.1016/S1473-3099(20)30120-1.

Countries which are not in John Hopkins' data have zero confirmed cases:

```{r zero-cases, eval = T}
pdat$cases[is.na(pdat$cases)] <- 0
```
### Model:

In the next chunk, we fit a linear regression model in which tourism is the IV and covid19 cases the DV (both logged):

```{r analysis, eval = T}
# Transform data to data.table format:
pdat <- data.table(pdat)

# Generate a few intermediary variables:
pdat[, LogCases := log(cases + 1)][, LogTourists := log((outbound_tour_groups_Q3_2019_improved + inbound_tour_groups_Q3_2019_1_improved + 1) / 2)]

# Fit a linear regression model on OECD countries:
NoPopModel <- lm(LogCases ~ LogTourists, pdat[oecd == TRUE])

# Print the summary of this model:
summary(NoPopModel)

# Generate predictions for all countries
pdat[, NoPopModPredictedLogCases := predict(NoPopModel, .SD)][, NoPopModPredictedCases := exp(NoPopModPredictedLogCases)][, NoPopModResidualLogCases := LogCases - NoPopModPredictedLogCases][, NoPopModResidualCases := cases - NoPopModPredictedCases]
```


```{r plots, eval = T}
ggplot(pdat[pdat$oecd, ], aes(y=cases, x=(inbound_tour_groups_Q3_2019_1_improved + outbound_tour_groups_Q3_2019_improved)/2))+
  geom_smooth(method = "lm")+
  scale_x_continuous(trans="log10")+
  scale_y_continuous(trans="pseudo_log")+
  geom_point(data=pdat[!pdat$oecd, ], aes(col="Other Countries"))+
  geom_point(aes(col = "OECD Countries"))+
  theme_minimal()+
  ylab("Confirmed covid19 cases\n(Log Scale)")+
  xlab("Tourism flows to and from China\n(In and Out-bound Tour Group Travellers, Log Scale)")+
  theme(legend.title = element_blank())

```

In models using data we cannot share here, we tested a range of other predictors, including flight passengers from Wuhan by country of destination in December, flights from Wuhan by destination country from October 2019 to April 2020 (scheduled), as well as covariates such as average temperature in January and February. Results were largely similar. We also tried alternative modelling strategies, including random forests, zero-inflated negative binomial models, and models which used per-capita measures of infection rates rather than the absolute number of confirmed cases, also with largely similar results. 


