---
title: "Graphic detail: The Chinese Century"
output: github_document
---

This is walk-through of the script required to reproduce the data and charts behind "The Chinese Century" published by The Economist, October 25th 2018

```{r}

#Set-up for srcipt 
#path <- "your path to our repo"
#setwd(path)

#Call in libraries
libs <- c("tidyverse", "WDI", "reshape2", "readxl", "countrycode")
lapply(libs, require, character.only=T)
```


We spent some time cleaning up historic GDP data from the Maddison Project Database, and merging that data with forecasts for world economic growth from the Economist Intelligence Unit. 
You can find Maddison data here : 
https://www.rug.nl/ggdc/historicaldevelopment/maddison/releases/maddison-project-database-2018

As you'll see, we don't have data for all years nor all countries, back to 1AD, but it's the best there is. 


```{r}
gdp <- read_csv("inputs/gdp.csv")
count(gdp, year)
```

We found the centre of each country, ie, it's latitude and longitude, via https://atcoordinates.info/resources/
We've assumed that a country's border hasn't changed over the years -- not true! -- but it's a good enough approximation for these purposes. 

```{r}
latlon.cou <- read.csv("inputs/country_centroids_all.csv", sep = "\t", stringsAsFactors = F) %>%
  select(iso2 = ISO3136, SHORT_NAME, LAT, LONG) %>%
  mutate(iso = countrycode(iso2, "iso2c", "iso3c")) %>%
  mutate(iso = ifelse(SHORT_NAME == "Namibia", "NAM", iso)) %>%
  bind_rows(., read.csv("inputs/missing_lat_longs.csv", stringsAsFactors=F)) %>% #add in missing lat~longs
  rename(cou.lon = LONG, cou.lat = LAT)
summary(latlon.cou)

```

Next we need to convert our spherical coordinates to Cartesian coordinates. 
h/t >>> https://rbrundritt.wordpress.com/2008/10/14/conversion-between-spherical-and-cartesian-coordinates-systems/)

```{r}

#Function 
earthRadius <- 1 #radius in km (6367)
conv.coord.f <- function(lat, lon) { 
  lat = lat / 180 * pi
  lon = lon / 180 * pi
  x = earthRadius * cos(lat)*cos(lon);
  y = earthRadius * cos(lat)*sin(lon);
  z = earthRadius * sin(lat);
  cart <- c(x, y, z) %>% t(.)
  return(cart) }

#apply function above, bind to df. 
cart.coord <- mapply(conv.coord.f, latlon.cou$cou.lat, latlon.cou$cou.lon) %>% t(.) %>% 
  as.data.frame(.) %>% rename(x=V1, y=V2, z=V3)
latlon.cou <- bind_cols(latlon.cou, cart.coord)
head(latlon.cou); tail(latlon.cou)

```


We add our Cartesian coordinates data to our main data frame, let's call it "DATA"
```{r}
DATA <- gdp %>%
  left_join(., select(latlon.cou, -iso2, -SHORT_NAME), by = "iso")
head(DATA); tail(DATA)
```


And then calcaulte the weighted average of our Cartesian coorindates. 
```{r}
eco.centre <- DATA %>% 
  group_by(year) %>% summarise(x = weighted.mean(x, gdp.pct.share, na.rm=T),
                               y = weighted.mean(y, gdp.pct.share, na.rm=T),
                               z = weighted.mean(z, gdp.pct.share, na.rm=T), 
                               count = n()) 
head(eco.centre); tail(eco.centre)
```


To plot the data on a map, we then need to convert our Cartesian coordinates back to spherical coordinates. 
```{r}
#function to convert cartesian to spherical coords (from same source above)
conv.caart.f <- function(x, y, z) {
  v <- ppls::normalize.vector(c(x, y, z))
  #v <- c(x, y, z)
  lat = asin(v[[3]]) / pi * 180
  lon = atan2(v[[2]], v[[1]]) / pi * 180
  latlon <- c(lat,lon)
  return(latlon) }

#apply function, bind
coord_centroid <- mapply(conv.caart.f, eco.centre$x, eco.centre$y, eco.centre$z) %>% t(.) %>% 
  as.data.frame(.) %>% rename(lat=V1, lon=V2)
eco.centre <- bind_cols(eco.centre, coord_centroid)
head(eco.centre); tail(eco.centre)

```


For our print version we plotted the data in QGIS using a Orthographic Projection centred on Moscow. 

But you can create a simple plot in R as follows: 

```{r}
#libs 
library(maps)
ggplot() +
  borders("world", colour="gray50", fill="gray50") +
  geom_point(data = eco.centre, aes(lon, lat)) +
  coord_fixed(1.3)

```

Write out the file. 

```{r}
write_csv(eco.centre, "eco_centre.csv")
```


